TGI-spinal-power

### 2022-05-03 analysis/tgispine_lme_power.Rmd
adding model of interest for cold location, and running power analysis

Create the first model. There is one model for each rating (instead of using rating as a parameter in the model).
First, going to try this on a very simple model where effect of manipulation (TGI vs. control) will be tested on rating of cold thermode, with subject as random effect
```{r}
# Summarise data per rating, per participant
# although summary not needed!
sum_pain <- aggregate(VASburning~ID*manipulation*condition*cold_probe, median, data = df_phs)
sum_warm <- aggregate(VASwarm~ID*manipulation*condition*cold_probe, median, data = df_phs)
sum_cold <- aggregate(VAScold~ID*manipulation*condition*cold_probe, median, data = df_phs)

# first try this on just one, simple model. The effect of manipulation on cold rating, with ID as a random effect
model.cold <- lmer(VAScold ~ manipulation + (1+manipulation|ID), data = df_phs)
summary(model.cold)
```
# View the coefficients for each person (should vary, due to random slope model)
```{r}
coef(model.cold)
```

>>>>

Create the first model. There is one model for each rating (instead of using rating as a parameter in the model).
First, going to try this on a very simple model where effect of manipulation (TGI vs. control) will be tested on rating of cold thermode, with subject as random effect
```{r}
# transform variables into proportions (aka divide by 100), this makes the effect size estimates more logical
df_phs$cold <- df_phs$VAScold/100
df_phs$warm <- df_phs$VASwarm/100
df_phs$burn <- df_phs$VASburning/100
df_phs$ID <- factor(df_phs$ID)

# Summarise data per rating, per participant
sum_burn <- aggregate(burn~ID*manipulation*condition*cold_location*procedure, median, data = df_phs)
sum_warm <- aggregate(warm~ID*manipulation*condition*cold_location*procedure, median, data = df_phs)
sum_cold <- aggregate(cold~ID*manipulation*condition*cold_location*procedure, median, data = df_phs)
# then compile all data again
sum_burn$VAS = 'burn'
names(sum_burn)[6] <- 'rating'
sum_warm$VAS = 'warm'
names(sum_warm)[6] <- 'rating'
sum_cold$VAS = 'cold'
names(sum_cold)[6] <- 'rating'
df_sum <- rbind(sum_cold, sum_warm)
df_sum <- rbind(df_sum, sum_burn)

# first try this on just one, simple model. The effect of manipulation on cold rating, with ID as a random effect
model.cold <- lmer(rating ~ manipulation + condition + cold_location +
                     (1|ID) + (1|procedure), 
                   data = df_sum[df_sum$VAS == 'cold' ,])
summary(model.cold)
```
# View the coefficients for each person (should vary, due to random slope model)
```{r}
coef(model.cold)
```

Create a model with more data points per variable, to extract more power
# First, let's try N = 100
```{r, include = FALSE}
# add more participants to the model, for simulation
model.coldN <- extend(model.cold, within = 'manipulation+condition+cold_location', n=100)
summary(model.coldN)
# check that number of data points per variable does = 30
dat_model <- xtabs(~ manipulation + condition + cold_location, data=getData(model.coldN))
print(dat_model)
```

# Run power analysis!
````{r}
# extract fixed effects from the model - take the lowest effect size from the model
tgi_effect <- fixef(model.coldN)["cold_locationprox_caud"]
# need to estimate minimum desired effect from the change in slope - considering values can range from 0-100, it might be simpler to think of these effect sizes as a proportion of 100 instead
# tgi_effect = -17.05 would perhaps leave an expected effect size of .17
# change the effect size to effect size of interest
fixef(model.cold)["manipulationTGI"] <- tgi_effect

# then run power analysis
powerSim(model.cold)
```

Now to run a power curve on the data to find out the optimal number of participants needed for the effect of interest
```{r}
# let's see if this works because it hasn't before
pc.cold <- powerCurve(model.coldN)
print(pc.cold)
```
